---
title: "myHomework2"
author: "chenpeng"
date: "2024-07-03"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: lumen
  pdf_document:
    latex_engine: xelatex
    toc: true
header-includes: 
  - \usepackage{ctex}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=40), tidy=TRUE)
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(DAAG)
library(dplyr)
library(MASS)
```


# 1. Loading and cleaning
a.
```{r}
ca_pa<-read.csv("../../data/calif_penn_2011.csv", header=T, encoding = "UTF-8")
```

b.
```{r}
nrow(ca_pa)
ncol(ca_pa)
```

c.
```{r}
colSums(apply(ca_pa,c(1,2),is.na))
```

Explanation:
it counts NA values in each column of ca_pa.

d.
```{r}
ca_pa.omit <- na.omit(ca_pa)
```

e.
```{r}
nrow(ca_pa.omit)
nrow(ca_pa) - nrow(ca_pa.omit)
```

f.
Answer: Yes. Because some rows have different type of NA value in columns.

# 2. This Very New House
a.
```{r}
ggplot(data = ca_pa.omit) +
    geom_point(aes(x =Built_2005_or_later , y = Median_house_value)) +
    labs(x = "Built 2005 or later",
       y = "Median house prices",
       title = "housing stock of California and Pennsylvania") +
    theme_bw()
```

b.
```{r}
ggplot(data = ca_pa.omit) +
    geom_point(aes(x =Built_2005_or_later , y = Median_house_value, color = as.factor(STATEFP))) +
    labs(x = "Built 2005 or later/%",
       y = "Median house prices",
       title = "housing stock of California and Pennsylvania",
       color = "State") +
    scale_color_manual(values = c("6" = "blue", "42" = "red"), labels = c("6" = "California", "42" = "Pennsylvania")) +
    theme_bw()
```

# 3. Nobody Home
a.
```{r}
ca_pa.omit$Vacancy_rate <- ca_pa.omit$Vacant_units/ca_pa.omit$Total_units

min_vacancy <- min(ca_pa.omit$Vacancy_rate)
max_vacancy <- max(ca_pa.omit$Vacancy_rate)
mean_vacancy <- mean(ca_pa.omit$Vacancy_rate)
median_vacancy <- median(ca_pa.omit$Vacancy_rate)

# Print the summary statistics
cat("Minimum Vacancy Rate:", min_vacancy, "\n")
cat("Maximum Vacancy Rate:", max_vacancy, "\n")
cat("Mean Vacancy Rate:", mean_vacancy, "\n")
cat("Median Vacancy Rate:", median_vacancy, "\n")
```

b.
```{r}
ggplot(data = ca_pa.omit) +
    geom_point(aes(x = Median_house_value, y = Vacancy_rate)) +
    labs(x = "Median house prices",
       y = "Vacancy rate",
       title = "housing stock of California and Pennsylvania") +
    theme_bw()
```

c.
```{r}
ggplot(data = ca_pa.omit) +
    geom_point(aes(x = Median_house_value, y = Vacancy_rate)) +
    labs(x = "Median house prices",
       y = "Vacancy rate",
       title = "housing stock of California and Pennsylvania") +
    facet_wrap(~ STATEFP, scales = "free", labeller = labeller(STATEFP = c("6" = "California", "42" = "Pennsylvania"))) +
    theme_bw()
```

Difference: 
In California, the higher the prices, the lower the Vacancy rate.
In Pennsylvania, the Median house prices are lower than those of California.

# 4.
a.
```{r}
acca <- c()
for (tract in 1:nrow(ca_pa.omit)) {
  if (ca_pa.omit$STATEFP[tract] == 6) {
    if (ca_pa.omit$COUNTYFP[tract] == 1) {
      acca <- c(acca, tract)
    }
  }
}
accamhv <- c()
for (tract in acca) {
  accamhv <- c(accamhv, ca_pa.omit[tract,10])
}
median(accamhv)
```

The block of code provided is intended to calculate the median house value for certain tracts within Alameda County (California) based on a dataset `ca_pa`. Let's break down how it achieves this:

1. **Initialization**:
   ```r
   acca <- c()
   ```
   - `acca` is initialized as an empty vector. It will store the indices (or row numbers) of the tracts that belong to Alameda County.

2. **First `for` loop**:
   ```r
   for (tract in 1:nrow(ca_pa)) {
     if (ca_pa$STATEFP[tract] == 6) {
       if (ca_pa$COUNTYFP[tract] == 1) {
         acca <- c(acca, tract)
       }
     }
   }
   ```
   - This loop iterates through each row (`tract`) of the dataset `ca_pa`.
   - It checks if the `STATEFP` value (state FIPS code) at the current row (`tract`) equals 6, which corresponds to California.
   - If the state is California (`STATEFP == 6`), it further checks if the `COUNTYFP` value (county FIPS code) at that row equals 1. County FIPS code 1 typically corresponds to Alameda County.
   - If both conditions are satisfied (tract is in California and specifically in Alameda County), it adds the index (`tract`) to the `acca` vector.

   **Purpose**: This loop identifies all the rows (tracts) in `ca_pa` that belong to Alameda County (California) and stores their indices in the `acca` vector.

3. **Second `for` loop**:
   ```r
   accamhv <- c()
   for (tract in acca) {
     accamhv <- c(accamhv, ca_pa[tract, 10])
   }
   ```
   - After identifying the tracts belonging to Alameda County, this loop iterates through each index (`tract`) in the `acca` vector.
   - For each index (`tract`), it retrieves the value in the 10th column (`ca_pa[tract, 10]`), which presumably represents the median house value for that tract.
   - It appends each median house value to the `accamhv` vector.

   **Purpose**: This loop extracts the median house values of all tracts within Alameda County from the dataset `ca_pa` and stores them in the `accamhv` vector.

4. **Calculate the median**:
   ```r
   median(accamhv)
   ```
   - Finally, `median(accamhv)` calculates the median of all the house values stored in the `accamhv` vector.

   **Purpose**: This computes and returns the median house value for all tracts within Alameda County.

**Summary**:
The entire block of code focuses on extracting and computing the median house value for tracts specifically within Alameda County (California) from the dataset `ca_pa`. It uses two loops: the first to identify the relevant tracts and the second to gather their median house values, culminating in the calculation of the median value itself.

b.
```{r}
median(ca_pa.omit$Median_house_value[ca_pa.omit$STATEFP == 6 & ca_pa.omit$COUNTYFP == 1])
```

c.
```{r}
mean.acc <- mean(ca_pa.omit$Built_2005_or_later[ca_pa.omit$STATEFP == 6 & ca_pa.omit$COUNTYFP == 1])
mean.sc <- mean(ca_pa.omit$Built_2005_or_later[ca_pa.omit$STATEFP == 6 & ca_pa.omit$COUNTYFP == 85])
mean.acp <- mean(ca_pa.omit$Built_2005_or_later[ca_pa.omit$STATEFP == 42 & ca_pa.omit$COUNTYFP == 3])

cat("average percentages of housing built since 2005:\n")
cat("Alameda County:",mean.acc,"\n")
cat("Santa Clara:",mean.sc,"\n")
cat("Allegheny County:",mean.acp,"\n")
```

d.
To compute correlations using the column names from your dataset `ca_pa`, you should modify the code accordingly. Here's how you can do it using the `names()` function to reference columns by name:

1. **Whole dataset**:
```{r}
cor(ca_pa.omit$Median_house_value, ca_pa.omit$Built_2005_or_later)
```

2. **California** (assuming STATEFP identifies California as 6):
```{r}
cor(ca_pa.omit$Median_house_value[ca_pa.omit$STATEFP == 6], ca_pa.omit$Built_2005_or_later[ca_pa.omit$STATEFP == 6])
```

3. **Pennsylvania** (assuming STATEFP identifies Pennsylvania as 42):
```{r}
cor(ca_pa.omit$Median_house_value[ca_pa.omit$STATEFP == 42], ca_pa.omit$Built_2005_or_later[ca_pa.omit$STATEFP == 42])
```

4. **Alameda County, California** (assuming COUNTYFP identifies Alameda County as 1 within STATEFP = 6):
```{r}
cor(ca_pa.omit$Median_house_value[ca_pa.omit$STATEFP == 6 & ca_pa.omit$COUNTYFP == 1], ca_pa.omit$Built_2005_or_later[ca_pa.omit$STATEFP == 6 & ca_pa.omit$COUNTYFP == 1])
```

5. **Santa Clara County, California** (assuming COUNTYFP identifies Santa Clara County as 85 within STATEFP = 6):
```{r}
cor(ca_pa.omit$Median_house_value[ca_pa.omit$STATEFP == 6 & ca_pa.omit$COUNTYFP == 85], ca_pa.omit$Built_2005_or_later[ca_pa.omit$STATEFP == 6 & ca_pa.omit$COUNTYFP == 85])
```

6. **Allegheny County, Pennsylvania** (assuming COUNTYFP identifies Allegheny County as 3 within STATEFP = 42):
```{r}
cor(ca_pa.omit$Median_house_value[ca_pa.omit$STATEFP == 42 & ca_pa.omit$COUNTYFP == 3], ca_pa.omit$Built_2005_or_later[ca_pa.omit$STATEFP == 42 & ca_pa.omit$COUNTYFP == 3])
```

Adjust the column names (`Median_house_value` and `Built_2005_or_later` in this example) according to the actual names in your dataset obtained from `names(ca_pa)`. This approach ensures that you're correctly referencing columns based on their names as listed in your dataset.

e.
```{r}
# Filter data for Alameda County, California (STATEFP = 6, COUNTYFP = 1)
alameda_data <- ca_pa.omit[ca_pa.omit$STATEFP == 6 & ca_pa.omit$COUNTYFP == 1, ]

# Filter data for Santa Clara County, California (STATEFP = 6, COUNTYFP = 85)
santa_clara_data <- ca_pa.omit[ca_pa.omit$STATEFP == 6 & ca_pa.omit$COUNTYFP == 85, ]

# Filter data for Allegheny County, Pennsylvania (STATEFP = 42, COUNTYFP = 3)
allegheny_data <- ca_pa.omit[ca_pa.omit$STATEFP == 42 & ca_pa.omit$COUNTYFP == 3, ]

# Create a combined plot
ggplot() +
  geom_point(data = alameda_data, aes(x = Median_household_income, y = Median_house_value, color = "Alameda County"), alpha = 0.7) +
  geom_point(data = santa_clara_data, aes(x = Median_household_income, y = Median_house_value, color = "Santa Clara County"), alpha = 0.7) +
  geom_point(data = allegheny_data, aes(x = Median_household_income, y = Median_house_value, color = "Allegheny County"), alpha = 0.7) +
  labs(x = "Median Household Income", y = "Median House Value", title = "Median House Value vs Median Household Income") +
  scale_color_manual(name = "County",
                     values = c("Alameda County" = "blue", "Santa Clara County" = "green", "Allegheny County" = "red")) +
  theme_minimal()
```

# MB.Ch1.11.
```{r}
gender <- factor(c(rep("female", 91), rep("male", 92)))
table(gender)
```
In this part, you created a factor variable gender with 91 "female" entries and 92 "male" entries.

```{r}
gender <- factor(gender, levels=c("male", "female"))
table(gender)
```
Here, you correctly reordered the levels of gender to prioritize "male" first and "female" second. Now the table shows 92 males and 91 females.



```{r}
gender <- factor(gender, levels=c("Male", "female"))
# Note the mistake: "Male" should be "male"
table(gender)
```
The mistake here is that you tried to relabel "male" as "Male" (with an uppercase 'M'), but since factors are case-sensitive in R, it created a new level "Male" which didn't exist previously. Hence, all entries previously labeled as "male" (lowercase 'm') are now labeled as <NA> because they don't match any of the specified levels.

```{r}
table(gender, exclude=NULL)
```
To see all levels including <NA>, you can use table(gender, exclude=NULL). This shows that there are 0 "Male", 91 "female", and 92 <NA> (missing) entries.


```{r}
rm(gender) # Remove gender
```
Finally, rm(gender) removes the gender variable from your workspace.

# MB.Ch1.12.
Let's proceed with writing a function that calculates the proportion of values in a vector `x` that exceed a specified `cutoff`.

### Part (a): Function Definition

Here's the function that calculates the proportion of values in a vector `x` that exceed a specified `cutoff`:

```{r}
proportion_exceed <- function(x, cutoff) {
  if (length(x) == 0) {
    stop("Vector x is empty.")
  }
  
  # Calculate proportion exceeding cutoff
  prop_exceed <- mean(x > cutoff)
  
  return(prop_exceed)
}
```

#### Explanation:
- **Function Name**: `proportion_exceed`
- **Parameters**:
  - `x`: Vector of numeric values.
  - `cutoff`: Numeric value against which elements of `x` are compared.
- **Returns**: Proportion of values in `x` that exceed `cutoff`.

### Testing with Sequence 1 to 100

Let's use the sequence of numbers 1 to 100 to verify our function:

```{r}
# Create sequence from 1 to 100
sequence <- 1:100

# Test the function with cutoff of 50
proportion_exceed(sequence, 50)
# Output: 0.5 (since half of the values in sequence exceed 50)
```

### Part (b): Applying to Escape Times Data

Now, we'll apply this function to the `ex01.36` dataset from the `Devore6` package, which contains escape times from an oil platform drill.

```{r}
# Load the Devore7 package (if not already installed, install it first)
# install.packages("Devore7")
library(Devore7)

# Load the dataset
data(ex01.36)

# View the dataset
head(ex01.36)

# Calculate the proportion of escape times that exceed 7 minutes (420 seconds)
prop_exceed_7min <- proportion_exceed(ex01.36, 420)

# Show the proportion
prop_exceed_7min
```

#### Explanation:
- **Loading Package**: `Devore6` package is loaded to access the `ex01.36` dataset.
- **Dataset**: We load the dataset and view its structure.
- **Calculation**: We calculate the proportion of escape times (`ex01.36`) that exceed 7 minutes (420 seconds).
- **Result**: `prop_exceed_7min` will give us the proportion of escape times exceeding 7 minutes.

### Visualization with `dotplot()`

To visualize the distribution of escape times, we can use `dotplot()`:

```{r}
# Visualize the distribution of escape times
dotplot(ex01.36$C1, main = "Distribution of Escape Times", xlab = "Escape Time", ylab = "data")
```

# MB.Ch1.18.
```{r}
# Load the Rabbit data
data(Rabbit)

Dose <- unstack(Rabbit, Dose ~ Animal)[,1]
Treatment <- unstack(Rabbit, Treatment ~ Animal)[,1]
BPchange <- unstack(Rabbit, BPchange ~ Animal)
data.frame(Treatment, Dose, BPchange)

```








