---
title: "myHomework1"
author: "chenpeng"
date: "2024-07-01"
output:
  html_document:
    toc: true
    toc_float: true
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 1
## T1
a.
```{r}
iowa.df<-read.csv("../../data/iowa.csv", sep = ';', header=T)
```

b.
```{r}
cat("iowa.df has",nrow(iowa.df),"rows and",ncol(iowa.df),"columns\n")
```

c.
```{r}
cat("The names of iowa.df:\n")
cat(names(iowa.df),sep = ", ")
```

d.
```{r}
cat("The value of row 5, column 7 of iowa.df:",iowa.df[5, 7])
```

e.
```{r}
cat("The second row of iowa.df:\n")
iowa.df[2, ]
```

## T2
a.
```{r,error=TRUE}
# 正确，向量赋值
vector1 <- c("5", "12", "7", "32")
# 正确，取最大值
max(vector1)
# 正确，字符串排序
sort(vector1)
# 错误，字符串不能求和
sum(vector1)
```

b.
```{r, error=TRUE}
vector2 <- c("5",7,12)
# 错误，由于此时vector2为字符型向量
vector2[2] + vector2[3]
```
```{r, error=TRUE}
# 正确，由于data.frame类型可以存储多种类型的数据
dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3]
```

```{r, error=TRUE}
list4 <- list(z1="6", z2=42, z3="49", z4=126)
# 正确，引用的是数值
list4[[2]]+list4[[4]]
# 错误，引用的是列表
list4[2]+list4[4]
```

## T3
a.
```{r}
seq(1, 1e4, by = 372)
seq(1, 1e4, length.out = 50)
```

b.
```{r}
# 1:3序列，重复输出整个序列三次
rep(1:3, times=3)
# 1:3序列，序列中每个数重复输出三次
rep(1:3, each=3)
```

## MB.Ch1.2
```{r}
library(DAAG)
obv <- DAAG::orings
obv_ch <- obv[c(1, 2, 4, 11, 13),]
plot(obv_ch$Temperature, obv_ch$Total,
     col = "blue",
     main = "Total Incidents vs. Temperature (Extracted Data)",
     xlab = "Temperature", ylab = "Incidents",
     xlim = c(min(obv_ch$Temperature)-1,max(obv_ch$Temperature)+1), ylim = c(min(obv_ch$Total)-1,max(obv_ch$Total)+1))
```

```{r}
plot(obv$Temperature, obv$Total,
     col = "blue",
     main = "Total Incidents vs. Temperature (Extracted Data)",
     xlab = "Temperature", ylab = "Incidents")
```

## MB.Ch1.4
(a)
```{r}
# Load the dataset 'ais'
data(ais)

# Use str() to get information about the dataset
str(ais)

# Check for missing values in each column
any_na <- apply(is.na(ais), 2, any)

# Print the result
cat("\n judge whether the column holds NA:\n")
print(any_na)
```

(b)
```{r}
gender_sport_table <- table(ais$sport, ais$sex)

gender_sport_df <- as.data.frame.matrix(gender_sport_table)

# Print the table
print(gender_sport_df)
```

## MB.Ch1.6
```{r}
lake_names <- c("Winnipeg", "Winnipegosis", "Manitoba", "SouthernIndian", 
                "Cedar", "Island", "Gods", "Cross", "Playgreen")
elevation <- c(217, 254, 248, 254, 253, 227, 178, 207, 217)
area <- c(24387, 5374, 4624, 2247, 1353, 1223, 1151, 755, 657)

Manitoba.lakes <- data.frame(elevation = elevation, area = area)
row.names(Manitoba.lakes) <- lake_names

print(Manitoba.lakes)
```

(a)
```{r}
attach(Manitoba.lakes)
plot(log2(area) ~ elevation, pch=16, xlim=c(170,280))
# NB: Doubling the area increases log2(area) by 1.0
text(log2(area) ~ elevation, labels=row.names(Manitoba.lakes), pos=4)
text(log2(area) ~ elevation, labels=area, pos=2)
title("Manitoba's Largest Lakes")
```

The label on the left of the point is its area.
The label on the right of the point is its name.
The y-axis is labeled as $\log_{2}(area)$. 

This transformation is applied because the area values of the lakes vary widely, with some being much larger than others. Using $\log_{2}(area)$ allows for better visualization by compressing the scale while maintaining the proportional differences between the lakes' areas. Each tick on the y-axis represents a doubling (or halving) of the area due to the logarithmic nature of the scale:

- A move upward by 1 unit on the y-axis corresponds to doubling the area of the lake.
- Conversely, a move downward by 1 unit on the y-axis corresponds to halving the area of the lake.

(b)
```{r (b)}
plot(area ~ elevation, pch=16, xlim=c(170,280), ylog=T)
text(area ~ elevation, labels=row.names(Manitoba.lakes), pos=4, ylog=T)
text(area ~ elevation, labels=area, pos=2, ylog=T)
title("Manitoba's Largest Lakes")
```